---
title: "Reproducible Document"
author: "Aylin S. Guller"
format: html
---

## Basics of R

```{r}
1 + 1
```

## About me 

- Southern Denmark university 
- **phd student**
- *in molecular biology*

## Simple Code

```{r}
3 * 3
```

## Testing of Git

cold 

## Setup code chunk 

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read.csv2(here::here("data/post-meal-insulin.csv"))

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)

tidier_post_meal_data <- small_post_meal_data |>
  rename(auc_glucose = auc_pg, 
         auc_insulin = auc_ins) |>
  filter( glykemi == 0) |>
  select(-glykemi)
```
## Showing the data

```{r}
post_meal_data
```

```{r}
glimpse(post_meal_data)
```

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

## Exercise: discrete plots

```{r}
#| fig-cap: "groups"
#| label: fig-group-bar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-group-bar above for a cool plot!

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = glykemi)) +
  geom_bar()
```

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```
## Putting two plots side by side 

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

```{r side-by-side-figs}
#| label: fig-bmi-group
#| fig-cap: "BMI and groups, side. by side"
#| fig-subcap:
#| - "Distribiton of BMI"
#| - "Groups"
#| layout-ncol: 2

ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```


## Plotting two continuous variables 

```{r}
#| fig-cap: "scatterplot of BMI versus AUC pf blood glucose "
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```
check out the @fig-bmi-auc-pg

## Pushing changes to remote 


## Pulling changes locally 

## Selecting data
```{r}
select(post_meal_data, Age)

select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group)

select(post_meal_data, starts_with("PG"))
select(post_meal_data, ends_with("DXA"))
select(post_meal_data, contains("0"))
```

##Renaming spesific columns
```{r}
small_post_meal_data |>
  rename(auc_glucose = auc_pg)

small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```

## Select and rename tasks

```{r}
# Task 1.
post_meal_data |>
  select(contains("OGTT"))

# Task 2.
post_meal_data |>
  select(contains(".."), -contains("...")) |>
  rename(
    P.Glucose.minus5.OGTT = P.Glucose..5.OGTT,
    Insulin.minus5.OGTT.X = Insulin..5.OGTT.X
  )

# Task 3.
basic_info <- select(post_meal_data, id, BMI, Length, Weight, Age)
rename(basic_info, Height = Length)

post_meal_data |>
  select(id, BMI, Length, Weight, Age) |>
  rename(Height = Length)

```


## filtering data by raw
```{r}
small_post_meal_data |>
  filter(glykemi == 1)

small_post_meal_data |>
  filter(BMI >=25)

small_post_meal_data |>
  filter(Group == "CTR")

```
```{r}
TRUE & TRUE 
```

```{r}
small_post_meal_data |>
  filter(Group =="CTR" , BMI >= 25)

small_post_meal_data |>
  filter( Group == "CTR" | BMI >=25)
```

